Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-20 paper/paper.tex
Changes to section 5.3

We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
experiments were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
due to the high parallelism of workflows, the resource utilization remains 
high most of the time, which leads to the rapid provisioning of new resources 
at the beginning of ensemble execution. This, in turn, results in less 
efficient total utilization due to high levels of parallelism. Setting the 
scaling limit to 1.0, ensures that the shape of the resource provisioning 
plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
of VMs before shutdown.

@@

There are multiple ways to evaluate the algorithms and also different metrics
are relevant to the application users that are needed to take the planning
decisions. The set of metrics for each ensemble run we prepared includes:
\begin{itemize}
  \item number of workflows completed $N_c$,
  \item user preference score (utility function) $s$,
  \item total amount of computing time completed $T_C$,
  \item effective cost of computing per hour $C_{eff}$, which includes
  overheads.
\end{itemize}

The number of workflows completed is a simplest metric and allows distinguishing
which algorithm performs better. However, as the workflows have
different priorities and sizes, a less efficient algorithm may schedule smaller
low-priority workflows first, thus increasing the total number of workflows
complete. It may also happen that too conservative admission algorithm may reject
overestimate the costs of some large workflows, thus reject some high-priority
ones which actually could fit into the schedule. Therefore we defined a user
preference score as:

%\begin{equation}
%\label{eq:score}
%s = \sum_{DAGs\ completed}(1/(p+1))
%\end{equation}

\begin{equation}
\label{eq:score}
s = \sum_{DAGs\ completed}2^{-p}
\end{equation}

where priorities of subsequent workflows in the ensemble are $p=0,1,2,\ldots$ (0
means highest priority). This exponential scoring function gives a strong weight
to the top priority workflows, since it has a property that a score for
completing a single workflow with priority $p$ is higher that the score for
completing all lower priority ones: 
\begin{equation}
\label{eq:score-property}
2^{-p} > \sum_{i=p+1,\ldots}2^{-i}
\end{equation}

Another metric is the total actual computing time completed and it is calculated
as the sum of all the runtimes of all the tasks of completed workflows. As in
our study we assumed that the resource cost pre hour is equal to \$1, the total
computing time can be easily compared to the allocated budget $b$. Their ratio
$u = (T_C * \$1)/b$ is the resource utilization and its inverse is the effective
resource cost: $C_{eff} = b/T_C$ in dollars per hour.

>>>>

We show the results for three algorithms: DPDS, WA-DPDS and SPSS. These 
experiments were run with maximum autoscaling factor set to 1.0 for
DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found that, 
due to the high parallelism of workflows, the resource utilization remains 
high most of the time, which leads to the rapid provisioning of new resources 
at the beginning of ensemble execution. This, in turn, results in less 
efficient total utilization due to high levels of parallelism. Setting the 
scaling limit to 1.0, ensures that the shape of the resource provisioning 
plan  is close to a rectangle (see Fig.~\ref{fig:spds-example}) and the VM 
termination procedure (Algorithm~\ref{alg:prov}) allows maximum utilization 
of VMs before shutdown. For the SPSS algorithm we set the $\alpha$ parameter
for deadline distribution to be 0.7.

@@

There are many different ways to evaluate and compare the performance of 
the algorithms. The set of ensemble performance metrics we have used to
compare our algorithms includes:

\begin{itemize}
  \item $N_c$: number of workflows completed,
  \item $s$: user preference score (utility function),
  \item $T_C$: total amount of computing time completed,
  \item $C_{eff}$: effective cost of computing per hour, which includes
  overheads.
\end{itemize}

The number of workflows completed is a simple metric that can be used for
comparison. However, because our ensembles contain workflows with different 
priorities and sizes, a less efficient algorithm may schedule smaller
workflows first, completing a large number of low-priority workflows. It may 
also happen that a conservative admission algorithm may overestimate the cost 
of a large workflow, thus rejecting high-priority workflows which actually 
could fit into the schedule. Therefore, we defined the user preference score as:

\begin{equation}
\label{eq:score}
s = \sum_{workflows\ completed}{2^{-d}}
\end{equation}

where $p$ is the priority of the workflow such that the highest-priority workflow 
has $p=0$, the next highest workflow has $p=1$, and so on. This exponential scoring 
function gives the highest priority workflow a score that is higher than all the 
lower-priority workflows combined:

\begin{equation}
\label{eq:score-property}
2^{-p} > \sum_{i=p+1,\ldots}2^{-i}
\end{equation}

This scoring is consistent with our definition of the problem, which is to 
complete as many workflows as possible, according to their priorities, given 
a budget and deadline.

The total amount computing time completed, $T_C$, is calculated as the sum 
of the runtimes of all tasks in all completed workflows. Because we assumed 
that the resource cost per hour is equal to \$1, the total computing time 
can be easily compared to the allocated budget $b$. Their ratio: 
$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour.
